Attribution of undesirable character traits, rather than trait-based dehumanization, predicts punishment decisions

### 2024-07-03 manuscript.Rmd
Update manuscript.Rmd

@Bastian2013 seek to account for this possibility by statistically controlling for participants’ moral outrage at the targets’ behaviour in their analysis.  They report that the relationship between trait-based dehumanization and punishment remains even when moral outrage is controlled for.  While this is interesting and suggestive of the independent effects of dehumanization, it cannot fully address the conceptual weaknesses in how dehumanization was operationalised.  A more convincing way to de-confound evidence for trait-based dehumanization from evidence of negative evaluation is to ask participants to rate the target group on traits that are uniquely human but vary from socially desirable to undesirable [@Over2021a, @Over2021b].  Previous research conducted by @Enock2021b has established that undesirable character traits such as jealous, spiteful and bitter are considered unique to humans and socially undesirable.  Across three intergroup contexts, the researchers found that participants attributed socially desirable human traits more strongly to the ingroup and socially undesirable traits more strongly to the outgroup; see also [@Decker2023; @Enock2021a; @Enock2022; @Enock2023].  @Enock2021b concluded that intergroup preference may better explain apparent evidence for trait-based dehumanization.  However, it is not yet clear how the attribution of uniquely human character traits relates to harm.  Addressing this question is crucial to understanding the extent to which the dual model of dehumanization can help explain real-world discrimination and negativity.

>>>>

@Bastian2013 seek to account for this possibility by statistically controlling for participants’ moral outrage at the targets’ behaviour in their analysis.  They report that the relationship between trait-based dehumanization and punishment remains even when moral outrage is controlled for.  While this is interesting and suggestive of the independent effects of dehumanization, it cannot fully address the conceptual weaknesses in how dehumanization was operationalised.  A more convincing way to de-confound evidence for trait-based dehumanization from evidence of negative evaluation is to ask participants to rate the target group on traits that are uniquely human but vary from socially desirable to undesirable [@Over2021a, @Over2021b].  Previous research conducted by @Enock2021b has established that undesirable character traits such as jealous, spiteful and bitter are considered unique to humans and socially undesirable.  Across three intergroup contexts, the researchers found that participants attributed socially desirable human traits more strongly to the ingroup and socially undesirable traits more strongly to the outgroup (see also @Decker2023; @Enock2021a; @Enock2022; @Enock2023).  @Enock2021b concluded that intergroup preference may better explain apparent evidence for trait-based dehumanization.  However, it is not yet clear how the attribution of uniquely human character traits relates to harm.  Addressing this question is crucial to understanding the extent to which the dual model of dehumanization can help explain real-world discrimination and negativity.
